\newpage
\section{Factorization Techniques}

Factorization is the process of writing a mathematical expression as a product 
of its factors. This is a fundamental technique in algebra used to simplify expressions, 
solve equations, and analyze functions.

\subsection{Common Factor}

Factor out the greatest common divisor (GCD) of all terms.

\textbf{Example:}

\[
  6x^2 + 9x = 3x(2x + 3)
\]

\subsection{Difference of Squares}

A difference of squares follows the identity:

\[
  a^2 - b^2 = (a - b)(a + b)
\]

\textbf{Example:}

\[
  x^2 - 16 = (x - 4)(x + 4)
\]

\subsection{Sum of Cubes:}

\[
  a^3 + b^3 = (a + b)(a^2 - ab + b^2)
\]

\subsection{Difference of Cubes:}

\[
  a^3 - b^3 = (a - b)(a^2 + ab + b^2)
\]

\textbf{Example:}

\[
  x^3 + 8 = (x + 2)(x^2 - 2x + 4)
\]

\subsection{Trinomial: Special Case \texorpdfstring{\(x^2 + bx + c\)}{x² + bx + c}}

This is the case where \(a = 1\). Find two numbers whose product is \(c\) and sum is \(b\).

\textbf{Example:}

\[
  x^2 + 5x + 6 = (x + 2)(x + 3)
\]

\subsection{Trinomial: General Form \texorpdfstring{\(ax^2 + bx + c\)}{ax² + bx + c}}

Like in the previous case we are going to use a similar procedure, but we start by
multiplying and diving by \(\frac{a}{a}\) and then proceed to do the rest. Look at the example

\begin{align*}
3x^2 -5x - 2 &= \frac{3(3x^2 -5x - 2)}{3} \\
             &= \frac{{(3x)}^2 -5(3x) - 6}{3}\\ 
             &= \frac{(3x- 6)(3x+1)}{3} \\
             &= (x- 2)(3x +1) \\
\end{align*}

\subsection{Perfect Square Trinomial}

These follow the identities:

\[
  {(a + b)}^2 = a^2 + 2ab + b^2
\]

\[
  {(a - b)}^2 = a^2 - 2ab + b^2
\]

\textbf{Example:}

\[
  x^2 + 6x + 9 = {(x + 3)}^2
\]

\subsection{Substitution}

Substitute a more complex expression with a single variable, factor, then back-substitute.

\textbf{Example:}

\[
  x^4 + 2x^2 + 1 \Rightarrow \text{Let } y = x^2 \Rightarrow y^2 + 2y + 1 = {(y + 1)}^2 
  \Rightarrow {(x^2 + 1)}^2
\]

\subsection{Rationalization of Radicals}

Rationalizing removes radicals from the denominator.

\textbf{Example (Single Radical):}

\[
  \frac{1}{\sqrt{2}} = \frac{\sqrt{2}}{2}
\]

\textbf{Example (Binomial):}

\[
  \frac{1}{\sqrt{3} + 1} = \frac{\sqrt{3} - 1}{(\sqrt{3} + 1)(\sqrt{3} - 1)} = \frac{\sqrt{3} - 1}{2}
\]

\subsection{Horner’s Method}

Used to divide a polynomial by a binomial of the form \((x - r)\).

\textbf{Steps:}

\begin{enumerate}

  \item Write coefficients of the polynomial \(P_1(x)\).

  \item Find the factors of the term with \(x^0\).
   
  \item Bring down the first coefficient.

  \item Multiply it by \(r\), add to next coefficient. Repeat until you get to 
        the term with \(x^0\).
  
  \item If the final addition gives zero then you have found a zero of the polynomial which  
        Now can be written in the following form \(P_2(x)(x - r)\). With \(P(x)\) being the 
        new polynomial with the coefficients which were computed during the process to get the zero.

\end{enumerate}

\textbf{Example:}

Given is \(P_1(x) = x^3 - 6x^2 + 11x - 6\).

The factors of -6 are \(\{\pm 1, \pm 2, \pm 3, \pm 6\}\). We use \(r = 1\):

\[
  \begin{array}{r|rrrr}
  1 & 1 & -6 & 11 & -6 \\
    &   & 1 & -5 & 6 \\
  \hline
    & 1 & -5 & 6 & 0 \\
  \end{array}
  \Rightarrow P_2(x) = (x^2 - 5x + 6)(x - 1)
\]

Now by repeating the process with \(r = 2\)

\[
  \begin{array}{r|rrr}
  2 & 1 & -5 & 6 \\
    &   &  2 & -6 \\
  \hline
    & 1 & -3 & 0 \\
  \end{array}
  \Rightarrow P_3(x) = (x - 3)(x - 2)(x - 1)
\]

The result final result is \((x - 3)(x - 2)(x - 1)\)

\subsection{Long Division of Polynomials}

Use the same algorithm as numerical long division.

\textbf{Example:}

\polylongdiv{x^3+x^2-1}{x-1}

with the result: \(x^2 +2x + 2 + \frac{1}{x-1}\)
