\newpage
\section{Recursion}

Recursion is term used to refer to a 'thing' define in terms
of itself. An example the factorial \(n! = (n - 1)!\) with \(0! = 1\) being the
base case where the recursion stops.

\subsection{From Recursive to Closed Form}

A sequence defined recursively can often be rewritten in an explicit or closed form.

\emph{recursive form} defines \( a_n \) based on previous terms (e.g., \( a_n = a_{n-1} + 2 \)).

\emph{closed form} expresses \( a_n \) directly in terms of \( n \), with no reference to previous terms.

\subsubsection*{General Method (when linear):}
\begin{enumerate}
    \item Calculate some terms and watch the difference or quotient between them
    maybe the difference is clear and you can build a geom. or arith. sequence directly.
    \item Identify the type (e.g., linear, homogeneous, constant coefficients).
    \item Solve the characteristic equation (if linear).
    \item Use initial conditions to determine constants.
\end{enumerate}

\subsection{Linear Recursion}

A linear recurrence has the form:

\begin{align*}
a_n & = \lambda a_{n -1} + c \\
a_1 & = \lambda a_0 + c \implies a_2 = \lambda (\lambda a_0 + c) + c \\
\implies a_n & = \lambda^n a_0 + (1 + \lambda + \lambda² + \cdots + \lambda^{n - 1})c
\end{align*}

Therefore

\[a_n = \lambda^n a_0 + \frac{\lambda^n - 1}{\lambda - 1}c\]

\subsection{Recursive Inequalities}

A \emph{recursive inequality} bounds a sequence rather than defines it exactly.
\\\\
\textbf{Example:} Suppose

\[
a_{n+1} \le \frac{1}{2} a_n + 3
\]

To analyze:
\begin{itemize}[label=\(-\)]
    \item Guess a bound (e.g., show \( a_n \le M \) for all \( n \))
    \item Use induction or iteration to justify convergence or boundedness
    \item Compare with a simpler sequence (e.g., geometric series)
\end{itemize}

These techniques are common in analysis, approximation, and numerical algorithms.

\subsection{Fibonacci Numbers and Closed Form}

The Fibonacci sequence is defined recursively:
\[
F_0 = 0, \quad F_1 = 1, \quad F_n = F_{n-1} + F_{n-2}
\]

The characteristic equation is:
\[
x^2 - x - 1 = 0 \Rightarrow x = \varphi = \frac{1 + \sqrt{5}}{2}, \quad \psi = \frac{1 - \sqrt{5}}{2}
\]

The closed form (Binet’s Formula) is:
\[
F_n = \frac{1}{\sqrt{5}} \left( \varphi^n - \psi^n \right)
\]

\begin{itemize}[label=\(-\)]
    \item \( \varphi \approx 1.618 \) is the golden ratio
    \item This formula demonstrates exponential growth of \( F_n \)
\end{itemize}

