\section{Introduction}

Hello my name is Miguel, this compendium is not mean to be a complete guide/book about all of mathematics,
but a collection of theorems, definitions, and proofs that I have found useful in my studies.
This compendium is not a replacement for a the books or the lectures at your university, but a complement to it.
On the one side I will try to keep it as simple as possible, but sometimes there are going to topics that will be quite 
complex and on the other side some proofs will be 
skiped or not included at all, because I think that they 
are not necessary for the understanding of the topic and may lead to confusion.
\\\\
My main idea while wirting this was to take as musch as I could from the books I have read, the lectures 
I have attended, the videos I have watched 
and the notes I have taken. This is also the reason why the Order may seem a bit strange, 
but I think that it is the best way to give an overiew about a lot of topics and also via the table of contents you can
easily find the topic you are looking for.
\\\\
The whole compendium is written in \LaTeX, so if you find any mistake or you want to add something, please 
feel free to do so in your own computer. If you see mistakes or you want to add something to the online version, please
let me know and I will try to fix it as soon as possible.
I will try to keep the compendium updated as much as 
I can, but I am not a professional writer nor a \LaTeX\ veteran, so please be patient with me.
\\\\
If you find this useful please consider maybe donating
to this project, but dont worry I this document to be free for ever.


\newpage