\newpage
\section{Linear Maps}

A linear map \( f: V \to W \) is a function that satisfies the following properties:

\begin{itemize}
    \item \( f(v_1 + v_2) = f(v_1) + f(v_2) \) for all \( v_1, v_2 \in V \).
    \item \( f(\lambda v) = \lambda f(v) \) for all \( v \in V \) and \( \lambda \in \Reals \).
    \item \( f(0) = 0 \).
\end{itemize} 

This kind of function is also called \emph{Homomorphism}. If \(V = W\) is it called \emph{Endomorphism}.

\subsection{Types of Linear Maps}

\begin{itemize}
    \item \emph{Injective or Monomorphism} (One-to-One): A linear map \( f: V \to W \) is injective if \( f(v_1) = f(v_2) \) implies \( v_1 = v_2 \).
    \item \emph{Surjective or Epimorphism} (Onto): A linear map \( f: V \to W \) is surjective if for every \( w \in W \), there exists a \( v \in V \) such that \( f(v) = w \).
    \item \emph{Bijective or Isomorphism}: A linear map \( f: V \to W \) is bijective if it is both injective and surjective.
\end{itemize}

\emph{Isomorphic} if there exists a bijective linear map between them. In this case, we can say 
that the two vector spaces are \emph{isomorphic}, 
and we write \( V \cong W \).

\subsection{Properties of Linear Maps}

\begin{itemize}
    \item The composition of two linear maps is a linear map.
    \item The inverse of a bijective linear map is also a linear map.
    \item The zero map \( f: V \to W \) defined by \( f(v) = 0 \) 
          for all \( v \in V \) is a linear map.
    \item The identity map \( \text{id}_V: V \to V \) defined by \( \text{id}_V(v) = v \) for all 
          \( v \in V \) is a linear map.
    \item The sum of two linear maps \( f: V \to W \) 
          and \( g: V \to W \) is a linear map defined by \( (f + g)(v) = f(v) + g(v) \).
    \item The scalar multiplication of a linear map \( f: V \to W \) by a scalar \( c \) is a linear 
          map defined by \( (cf)(v) = c(f(v)) \).
    \item The composition of linear maps is associative, i.e., \( (f \circ g) \circ h = f \circ (g \circ h) \).
    \item The composition of linear maps is distributive over addition, i.e., 
          \( f \circ (g + h) = f \circ g + f \circ h \).
    \item The composition of linear maps is compatible with scalar multiplication, i.e., 
          \( (cf) \circ g = c(f \circ g) \).
    \item Linear Maps compose a vector space 
          over the field of scalars.
          
          \[
            (Hom(V,W), +, \cdot )
          \]
\end{itemize}

\subsection{The Kernel of a Linear Map}

The kernel of a linear map \( f: V \to W \) is the set of all vectors in \( V \) that are mapped to 
the zero vector in \( W \):

\[
    \ker(f) = \{ v \in V \mid f(v) = 0 \}
\]

\subsection{The Image of a Linear Map}

The image of a linear map \( f: V \to W \) is the set of all vectors in \( W \) that can be expressed 
as \( f(v) \) for some \( v \in V \):

\[
    \text{Im}(f) = \{ w \in W \mid w = f(v) \text{ for some } v \in V \}
\]
 
\subsection{The Rank of a Linear Map}

The rank of a linear map \( f: V \to W \) is the dimension of its image:

\[
    \text{rank}(f) = (\text{Im}(f))
\]

\subsection{The Nullity of a Linear Map}

The nullity of a linear map \( f: V \to W \) is the dimension of its kernel:
    
\[
    \text{nullity}(f) = (\ker(f))
    
\]

\subsection{The Rank-Nullity Theorem}

The rank-nullity theorem states that for a linear map \( f: V \to W \):

\[
    \text{}(\ker(f)) + \text{dim}(\text{Im}(f)) = \text{dim}(V).
\]

\subsection{Proof of the injectivity of the Kernel}

Suppose we have a linear map \emph{f} and two vectors \(v_1\) and \(v_2\) which are not equal. 
We are going to assume that they both map to the \(\vec{0}\) therefore:

\[
    f(v_1) = f(v_2) = \vec{0}
\]

Because of the linearity of the map we can write:

\[
    f(v_1 - v_2) = f(v_1) - f(v_2) = \vec{0} - \vec{0} = \vec{0}
\]

And that contradicts the assumption that \(v_1\) and \(v_2\) are not equal. Therefore, the 
kernel of a linear map is injective.

\subsection{Dimension Formula for Linear Mappings}

Let \(f: V \to W\) be linear and \(\dim(V) = n\). Then
    
\[
    \dim(\ker(f)) + \operatorname{rank}(f) = n.
\]
    
Remember that \(\ker(f)\) forms a subspace of \(V\) and therefore, \(\dim(\ker(f)) := r \leq n\). We extend an arbitrary basis \((v_1, \ldots, v_r)\) of \(\ker(f)\) to a basis \((v_1, \ldots, v_r, v_{r+1}, \ldots, v_n)\) of \(V\). Setting \(w_{r+i} = f(v_{r+i})\) for \(i = 1, \ldots, n - r\), we have \(\forall v \in V\):

\begin{align*}
    f(v) &= f(\lambda_1v_1 + \cdots + \lambda_r v_r + \lambda_{r+1}v_{r+1} + \cdots + \lambda_n v_n) \\
    &= \lambda_1 \underbrace{f(v_1)}_{=0} + \cdots + \lambda_r \underbrace{f(v_r)}_{=0} + \lambda_{r+1} f(v_{r+1}) + \cdots + \lambda_n f(v_n) \\
    &= \lambda_{r+1} f(v_{r+1}) + \cdots + \lambda_n f(v_n) \\
    &= \lambda_{r+1}w_{r+1} + \cdots + \lambda_n w_n
\end{align*}
    
Thus, \(\operatorname{Im}(f) = \operatorname{span}(w_{r+1}, \ldots, w_n)\). We now show that \(w_{r+1}, \ldots, w_n\) are linearly independent. Let 
    
\[
    \lambda_{r+1}w_{r+1} + \cdots + \lambda_n w_n = 0
\]
    
From

\[
    0 = \lambda_{r+1}w_{r+1} + \cdots + \lambda_n w_n = f(\lambda_{r+1}v_{r+1} + \cdots + \lambda_n v_n)
\]
    
it follows that

\[
    \lambda_{r+1}v_{r+1} + \cdots + \lambda_n v_n \in \ker(f)
\]
    
Therefore,
    
\[
    \lambda_{r+1}v_{r+1} + \cdots + \lambda_n v_n = \lambda_1v_1 + \cdots + \lambda_r v_r
\]

For some \(\lambda_1, \ldots, \lambda_r\). Since \(v_1, \ldots, v_n\) are linearly independent, 
we have \(\lambda_1 = \cdots = \lambda_n = 0\). Thus, the vectors \(w_{r+1}, \ldots, w_n\) are linearly 
independent.
    
It follows that \(\dim(\operatorname{Im}(f)) = \operatorname{rank}(f) = n - r\) and therefore,
    
\[
    \dim(\ker(f)) + \dim(\operatorname{Im}(f)) = \dim(\ker(f)) + \operatorname{rank}(f) = r + n - r = n
\]

\QED
  
\subsection{Identifying the type of linear map}

\begin{itemize} 
    \item If \(\dim(\ker(f)) = 0\) then the map is injective.
    \item If \(\dim(\ker(f)) = \dim(V)\) then the map is surjective.
    \item If \(\dim(\ker(f)) = \dim(V)\) and \(\dim(W) = 0\) then the map is bijective.
\end{itemize}

\subsection{Image of the basis}

Let \((V) = \dim(W)\), For the basis \((v_1, \dots, v_n)\) of \(V\) we have the image \((w_1, \dots, w_n)\) of \(W\). Then:

\begin{itemize}
    \item If \(\dim(\ker(f)) = 0\) then the map is injective and \((w_1, \dots, w_n)\) is a basis of \(W\).    
    \item If \(\dim(\ker(f)) = \dim(V)\) then the map is surjective and \((w_1, \dots, w_n)\) is a spanning set of \(W\).
    \item If \(\dim(\ker(f)) = \dim(V)\) and \(\dim(W) = 0\) then the map is bijective and \((w_1, \dots, w_n)\) is a 
    basis of \(W\).
\end{itemize}

\subsection{Linear Maps and Matrices}

Let \( V \) and \( W \) be finite-dimensional vector spaces over the same field \( K \). 
If \( \dim V = n \) and \( \dim W = m \), then a linear map \( f: V \to W \) can be represented by 
an \( m \times n \) matrix. The action of the linear map on a vector can be expressed as:

\[
    f(x) = A \cdot x,
\]

where \emph{A} is the matrix representation of \emph{f} and \( v \) is the vector represented in a column 
format. The columns of the matrix \emph{A} are the images of the basis vectors of \( V \) under the linear 
map \emph{f}.

