\newpage
\section{Topology}

In this section, we introduce essential vocabulary used in topology. Each term is accompanied by a 
brief explanation and its formal mathematical definition.

\subsection{Introduction to topological nomenclature}

\subsubsection{Open Set}
	     
A subset \( U \subseteq X \) of a topological space is called 
\emph{open} if for every point \( x \in U \), there exists an \( \varepsilon > 0 \) 
such that the open ball \( B_\varepsilon(x) \subseteq U \). 
Intuitively, an open set contains none of its boundary points and every point has some 
wiggle room around it.

\subsubsection{Closed Set} 
	      
A subset \( A \subseteq X \) is called \emph{closed} if its 
complement \( X \setminus A \) is open. Equivalently, \emph{A} contains all its limit points. 
That is, \emph{A} is closed if it includes its boundary.

\subsubsection{Interior Point}

A point \( x \in A \) is an \emph{interior point} of \( A \subseteq X \) if there 
exists \( \varepsilon > 0 \) such that \( B_\varepsilon(x) \subseteq A \). 
The set of all interior points of \emph{A} is called the \emph{interior} of \emph{A}, 
denoted \( \mathrm{int}(A) \).

\subsubsection{Boundary Point}
	      
A point \( x \in X \) is a \emph{boundary point} of a set \( A \subseteq X \) 
if every open ball around \emph{x} contains both points in \emph{A} and in \( X \setminus A \). 
The set of all boundary points is called the \emph{boundary} of \emph{A}, denoted \( \partial A \).

\subsubsection{Accumulation Point / Limit Point}

A point \( x \in X \) is an \emph{accumulation point} of a set \( A \subseteq X \)
if every open ball \( B_\varepsilon(x) \) contains a point of \( A \setminus \{x\} \). 
In other words, points of \emph{A} cluster arbitrarily close to 
\emph{x}, even if \( x \notin A \).

\subsubsection{Isolated Point} 
	      
A point \( x \in A \) is an \emph{isolated point} if there exists \( \varepsilon > 0 \)
such that \( B_\varepsilon(x) \cap A = \{x\} \). 
That is, \emph{x} stands alone in \emph{A} without other points of \emph{A} nearby.

\subsubsection{Compact Set} 

A set \( K \subseteq X \) is \emph{compact} if every open cover of \( K \) has a finite sub-cover. 
In \(\Reals^n\), this is equivalent to \( K \) being closed and bounded (by the Heine–Borel theorem).

\subsubsection{Dense Set} 
	      
A subset \( D \subseteq X \) is \emph{dense} in \emph{X} if every point 
\( x \in X \) is either in \( D \) or is a limit point of \( D \). 
Equivalently, the closure of \( D \) is \emph{X}, i.e., \( \overline{D} = X \).

\subsubsection{Open Ball} (\( B_\varepsilon(x) \)) 

For a metric space \( (X, d) \), the \emph{open ball} centered at 
\( x \in X \) with radius \( \varepsilon > 0 \) is defined as: 
	      
\[
	B_\varepsilon(x) := \{ y \in X \mid d(x, y) < \varepsilon \}
\]

It represents the set of all points within distance \( \varepsilon \) from \emph{x}, excluding the boundary.
